---
title: "Now and Later R Analyses Revision 1"
author: "RSF"
date: "10/11/2023"
output: 
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1: Set-up
```{r}
setwd("~/Desktop/MONITOR_Now_Later/allquery")
now<-read.table('Now/combinedfeatquery.txt')
later<-read.table('Later/combinedfeatquery.txt')

now_means<-subset(now, V2=="stats/cope1")
later_means<-subset(later, V2=="stats/cope1")


Now_pcnt_means<-now_means$V6
Later_pcnt_means<-later_means$V6

ID<-c("MO-001","MO-009","MO-023","MO-030","MO-033","MO-035","MO-063","MO-078","MO-095","MO-101","MO-106",
      "MO-109","MO-113","MO-127","MO-138","MO-141","MO-151","MO-153","MO-159","MO-170","MO-188","MO-190",
      "MO-205","MO-209","MO-223","MO-276")

copes<-as.data.frame(cbind(ID,Now_pcnt_means,Later_pcnt_means))

setwd("~/Desktop/Manuscripts/Ongoing Projects/MONITOR Now and Later/Data")
library(openxlsx)
MONITOR<-read.xlsx("Monitor Full Data Set REVISED.xlsx")

MONITOR_2<-subset(MONITOR, ID=="MO-001" | ID=="MO-009" | ID=="MO-023" | ID=="MO-030" | ID=="MO-033" |
                    ID=="MO-035" | ID=="MO-063" | ID=="MO-078" | ID=="MO-095" | ID=="MO-101" | ID=="MO-106" |
                    ID=="MO-109" | ID=="MO-113" | ID=="MO-127" | ID=="MO-138" | ID=="MO-141" | ID=="MO-151" | 
                    ID=="MO-153" | ID=="MO-159" | ID=="MO-170" | ID=="MO-188" | ID=="MO-190" | ID=="MO-205" |
                    ID=="MO-209" | ID=="MO-223" | ID=="MO-276")

data<-merge(MONITOR_2,copes, by="ID")
data<-data[c(1:3,87,88,6:8,10,16,17,19,21,22,23,24,26,27,45,46,79,80,82:85,20,86)]
fulldata<-MONITOR[c(1:3,6:8,10,16,17,19,21,22,23,24,26,27,45,46,79,80,82:85,20,86)]

#Full Data#
fullvarying <- fulldata[c(1:2,4,10:26:length(fulldata))]
fullbaseline <- subset(fulldata[-c(4,10:26:length(fulldata))],Timepoint==1)

fullvarying.1 <- subset(fullvarying,Timepoint==1)
fullvarying.2 <- subset(fullvarying,Timepoint==2)
fullvarying.3 <- subset(fullvarying,Timepoint==3)
fullvarying.4 <- subset(fullvarying,Timepoint==4)

colnames(fullvarying.1) <- paste(colnames(fullvarying.1),"1",sep=".")
colnames(fullvarying.2) <- paste(colnames(fullvarying.2),"2",sep=".")
colnames(fullvarying.3) <- paste(colnames(fullvarying.3),"3",sep=".")
colnames(fullvarying.4) <- paste(colnames(fullvarying.4),"4",sep=".")

library(dplyr)
fullwide.data <- left_join(fullbaseline,fullvarying.1,by=c("ID"="ID.1")) %>% 
  left_join(.,fullvarying.2,by=c("ID"="ID.2")) %>% 
  left_join(.,fullvarying.3,by=c("ID"="ID.3")) %>%
  left_join(.,fullvarying.4,by=c("ID"="ID.4"))
detach("package:dplyr", unload = TRUE)

#MRI Participant Data Set#
data$Now_pcnt_means<-as.numeric(data$Now_pcnt_means)
data$Later_pcnt_means<-as.numeric(data$Later_pcnt_means)

varying <- data[c(1:2,6,12:28:length(data))]
baseline <- subset(data[-c(6,12:26:length(data))],Timepoint==1)

varying.1 <- subset(varying,Timepoint==1)
varying.2 <- subset(varying,Timepoint==2)
varying.3 <- subset(varying,Timepoint==3)
varying.4 <- subset(varying,Timepoint==4)

colnames(varying.1) <- paste(colnames(varying.1),"1",sep=".")
colnames(varying.2) <- paste(colnames(varying.2),"2",sep=".")
colnames(varying.3) <- paste(colnames(varying.3),"3",sep=".")
colnames(varying.4) <- paste(colnames(varying.4),"4",sep=".")

library(dplyr)
wide.data <- left_join(baseline,varying.1,by=c("ID"="ID.1")) %>% 
  left_join(.,varying.2,by=c("ID"="ID.2")) %>% 
  left_join(.,varying.3,by=c("ID"="ID.3")) %>%
  left_join(.,varying.4,by=c("ID"="ID.4"))
detach("package:dplyr", unload = TRUE)

Timevars <- grep("Timepoint",colnames(wide.data),value=TRUE)
wide.data <- wide.data[,!(colnames(wide.data)%in%Timevars)]
colnames(wide.data) <- (gsub("_","",colnames(wide.data)))
colnames(wide.data) <- (gsub(".1","baseline",colnames(wide.data)))

library(plyr)
wide.data2<- rename(wide.data, c("WM.AgeCorrectedScorebaseline"="WMAgeCorrectedScorebaseline","WM.AgeCorrectedScore.2"="WMAgeCorrectedScore.2",
                                 "WM.AgeCorrectedScore.3"="WMAgeCorrectedScore.3", "WM.AgeCorrectedScore.4"="WMAgeCorrectedScore.4",
                                 "PSM.AgeCorrectedScorebaseline"="PSMAgeCorrectedScorebaseline","PSM.AgeCorrectedScore.2"="PSMAgeCorrectedScore.2",
                                 "PSM.AgeCorrectedScore.3"="PSMAgeCorrectedScore.3", "PSM.AgeCorrectedScore.4"="PSMAgeCorrectedScore.4",
                                 "CHAMPS.TotalPA.Hrsbaseline"="CHAMPSTotalPAHrsbaseline","CHAMPS.TotalPA.Hrs.2"="CHAMPSTotalPAHrs.2",
                                 "CHAMPS.TotalPA.Hrs.3"="CHAMPSTotalPAHrs.3","CHAMPS.TotalPA.Hrs.4"="CHAMPSTotalPAHrs.4","Champs.Total.SB.Hrsbaseline"="ChampsTotalSBHrsbaseline",
                                 "Champs.Total.SB.Hrs.2"="ChampsTotalSBHrs.2","Champs.Total.SB.Hrs.3"="ChampsTotalSBHrs.3","Champs.Total.SB.Hrs.4"="ChampsTotalSBHrs.4",
                                 "Champs.Frequencybaseline"="ChampsFrequencybaseline", "Champs.Frequency.2"="ChampsFrequency.2", "Champs.Frequency.3"="ChampsFrequency.3",
                                 "Champs.Frequency.4"="ChampsFrequency.4","Champs.MVPA.Frequencybaseline"="ChampsMVPAFrequencybaseline","Champs.MVPA.Frequency.2"="ChampsMVPAFrequency.2",
                                 "Champs.MVPA.Frequency.3"="ChampsMVPAFrequency.3", "Champs.MVPA.Frequency.4"="ChampsMVPAFrequency.4"))
detach("package:plyr", unload = TRUE)
wide.data2$time<-1
wide.data2<-wide.data2[c(1,82,2,5:81)]
data2 <- reshape(as.data.frame(wide.data2),idvar="ID",varying=c(27:80),direction="long",sep=".") 

library(dplyr)
wide.data3 <- left_join(baseline,varying.1,by=c("ID"="ID.1")) %>% 
  left_join(.,varying.2,by=c("ID"="ID.2")) %>% 
  left_join(.,varying.3,by=c("ID"="ID.3")) %>%
  left_join(.,varying.4,by=c("ID"="ID.4"))
detach("package:dplyr", unload = TRUE)

Timevars <- grep("Timepoint",colnames(wide.data3),value=TRUE)
wide.data3 <- wide.data3[,!(colnames(wide.data3)%in%Timevars)]
colnames(wide.data3) <- (gsub("_","",colnames(wide.data3)))

library(plyr)
wide.data3<- rename(wide.data3, c("WM.AgeCorrectedScore.1"="WMAgeCorrectedScore.1","WM.AgeCorrectedScore.2"="WMAgeCorrectedScore.2",
                                 "WM.AgeCorrectedScore.3"="WMAgeCorrectedScore.3", "WM.AgeCorrectedScore.4"="WMAgeCorrectedScore.4",
                                 "PSM.AgeCorrectedScore.1"="PSMAgeCorrectedScore.1","PSM.AgeCorrectedScore.2"="PSMAgeCorrectedScore.2",
                                 "PSM.AgeCorrectedScore.3"="PSMAgeCorrectedScore.3", "PSM.AgeCorrectedScore.4"="PSMAgeCorrectedScore.4",
                                 "CHAMPS.TotalPA.Hrs.1"="CHAMPSTotalPAHrs.1","CHAMPS.TotalPA.Hrs.2"="CHAMPSTotalPAHrs.2",
                                 "CHAMPS.TotalPA.Hrs.3"="CHAMPSTotalPAHrs.3","CHAMPS.TotalPA.Hrs.4"="CHAMPSTotalPAHrs.4","Champs.Total.SB.Hrs.1"="ChampsTotalSBHrs.1",
                                 "Champs.Total.SB.Hrs.2"="ChampsTotalSBHrs.2","Champs.Total.SB.Hrs.3"="ChampsTotalSBHrs.3","Champs.Total.SB.Hrs.4"="ChampsTotalSBHrs.4",
                                 "Champs.Frequency.1"="ChampsFrequency.1", "Champs.Frequency.2"="ChampsFrequency.2", "Champs.Frequency.3"="ChampsFrequency.3",
                                 "Champs.Frequency.4"="ChampsFrequency.4","Champs.MVPA.Frequency.1"="ChampsMVPAFrequency.1","Champs.MVPA.Frequency.2"="ChampsMVPAFrequency.2",
                                 "Champs.MVPA.Frequency.3"="ChampsMVPAFrequency.3", "Champs.MVPA.Frequency.4"="ChampsMVPAFrequency.4"))
detach("package:plyr", unload = TRUE)
wide.data3$time<-1
wide.data3<-wide.data3[c(1,82,5:81)]
data3 <- reshape(as.data.frame(wide.data3),idvar="ID",varying=c(8:79),direction="long",sep=".") 

data3$timefactor<-as.factor(data3$time)

data3$Group<-NA
data3$Group[data3$Treatment.Group==1]<-"Immediate Intervention"
data3$Group[data3$Treatment.Group==2]<-"Delayed Intervention"

data3$Timepoint<-NA
data3$Timepoint[data3$time==1]<-"Baseline"
data3$Timepoint[data3$time==2]<-"2-Months"
data3$Timepoint[data3$time==3]<-"4-Months"
data3$Timepoint[data3$time==4]<-"6-Months"
```


#2: Table 1. Participant Characteristics at Baseline

MRI Participants Only
```{r}
library(tableone)
vars<-dput(names(wide.data2[c(4:6,9,7,15,16,26)]))
Table1<-CreateTableOne(vars=vars, data=wide.data2)
print(Table1,contDigits=2, missing=TRUE, quote=FALSE)

Table1_strata<-CreateTableOne(vars=vars, strata = "Treatment.Group", data=wide.data2)
print(Table1_strata,contDigits=2, missing=TRUE, quote=FALSE)
```

Full data
```{r}
fullvars<-dput(names(fullwide.data[c(4:6,10,7,16,12,27)]))
Table1full<-CreateTableOne(vars=fullvars, data=fullwide.data)
print(Table1full,contDigits=2, missing=TRUE, quote=FALSE)

Table1full_strata<-CreateTableOne(vars=fullvars, strata = "Treatment.Group", data=fullwide.data)
print(Table1full_strata,contDigits=2, missing=TRUE, quote=FALSE)
```

#3: Figure 3. Group level changes in MVPA and SB over time
```{r}
library(lme4)
library(robustlmm)
library(lsmeans)
library(lmerTest)
library(ggplot2)
library(forcats)

#Used linear mixed model (controlling for baseline) to determine estimated marginal
#means for MVPA and SB by treatment group
emm_options(opt.digits = FALSE)
#Changes in MVPA over time
MVPA.change<-lmer(MVPAMin~factor(time)*Treatment.Group + (1|ID) + MVPAMinbaseline,data2)
anova(MVPA.change)
lsmeans(MVPA.change, ~Treatment.Group|time)
contrast(lsmeans(MVPA.change, ~Treatment.Group|time), "trt.vs.ctrl", adj="none")
confint(contrast(lsmeans(MVPA.change, ~Treatment.Group|time), "trt.vs.ctrl", adj="none"))

#Changes in SB over time
SB.change<-lmer(SBMin~factor(time)*Treatment.Group + (1|ID) + SBMinbaseline,data2)
lsmeans(SB.change, ~Treatment.Group|time)
contrast(lsmeans(SB.change, ~Treatment.Group|time), "trt.vs.ctrl", adj="none")
confint(contrast(lsmeans(SB.change, ~Treatment.Group|time), "trt.vs.ctrl", adj="none"))

#Create figure using lsmeans
MVPA.mean<-c(80.96,65.41,94.96,62.70,75.00,54.09,90.07,69.66) #MVPA means
MVPA.se<-c(13.9,14.6,9.4,9.8,9.7,9.8,9.7,9.8) #MVPA standard errors
SB.mean<-c(657.53,723.81,679.58,703.05,711.02,689.96,687.83,694.14) #SB means
SB.se<-c(30.30,37.78,21.01,21.81,21.68,21.81,21.67,21.81) #SB standard errors
tx <- c("I-INT", "D-INT", "I-INT", "D-INT", "I-INT", "D-INT", "I-INT", "D-INT") #Treatment Groups
time <- c("Baseline", "Baseline", "2-Months", "2-Months", "4-Months", "4-Months", "6-Months", "6-Months") #Timepoints
Fig2data<- data.frame(MVPA.mean, MVPA.se, SB.mean, SB.se, tx, time) #Combine in a dataframe

#MVPA Plot
library(plyr)
MVPA.group.graph<- Fig2data %>%
  mutate(time = fct_relevel(time, 
                            "Baseline", "2-Months", "4-Months", "6-Months")) %>%
  ggplot(aes(fill=tx, y=MVPA.mean, x=time)) + geom_bar(position=position_dodge(1), stat="identity") +
  geom_errorbar(aes(ymin = MVPA.mean - MVPA.se, ymax = MVPA.mean + MVPA.se), position=position_dodge(1), width = 0.2, stat="identity") +
  labs(x="", y="Moderate-to-Vigorous Physical Activity
(minutes/day)") +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + ylim(0,120)

MVPA.group.graph

#SB Plot
SB.group.graph<- Fig2data %>%
  mutate(time = fct_relevel(time, 
                            "Baseline", "2-Months", "4-Months", "6-Months")) %>%
  ggplot(aes(fill=tx, y=SB.mean, x=time)) + geom_bar(position=position_dodge(1), stat="identity") +
  geom_errorbar(aes(ymin = SB.mean - SB.se, ymax = SB.mean + SB.se), position=position_dodge(1), width = 0.2, stat="identity") +
  labs(x="", y="Sedentary Behaviour
(minutes/day)") +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + ylim(0,800)

SB.group.graph


detach("package:plyr", unload = TRUE)

```

#3: Figure 4: Individual changes in MVPA and SB Across Time
```{r}

library(lme4)
library(lsmeans)
library(forcats)
library(ggplot2)
library(plyr)

MVPA.Compliance.Graph<- data3 %>%
  mutate(Timepoint = fct_relevel(Timepoint, 
                            "Baseline", "2-Months", "4-Months", "6-Months")) %>%
ggplot() + geom_line(aes(x=Timepoint, y=MVPAMin, group=ID, linetype= Group)) +
labs(x="", y="Moderate-to-Vigorous Physical Activity 
  (minutes/day)") +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
MVPA.Compliance.Graph

SB.Compliance.Graph<- data3 %>%
  mutate(Timepoint = fct_relevel(Timepoint, 
                                 "Baseline", "2-Months", "4-Months", "6-Months")) %>%
  ggplot() + geom_line(aes(x=Timepoint, y=SBMin, group=ID, linetype= Group)) +
  labs(x="", y="Sedentary Behaviour 
(minutes/day)") +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + ylim(400,1100)
SB.Compliance.Graph
```


#4: Create data frames for FSL analyses
```{r}
setwd("~/Desktop/MONITOR_Now_Later")

###Cross-sectional Data###
baselineMVPA<-wide.data2[c(1,3,4,5,6,7,9,11,15,26)]
baselineMVPA$Agescale<-scale(baselineMVPA$Age, scale=FALSE)
baselineMVPA$BMIscale<-scale(baselineMVPA$BMIbaseline, scale=FALSE)
baselineMVPA$MOCAscale<-scale(baselineMVPA$MoCA, scale=FALSE)
baselineMVPA$MVPAscale<-scale(baselineMVPA$MVPAMinbaseline, scale=FALSE)
baselineMVPA$SBscale<-scale(baselineMVPA$SBMinbaseline, scale=FALSE)
baselineMVPA$SBscale<-scale(baselineMVPA$SBMinbaseline, scale=FALSE)
baselineMVPA$Gender[baselineMVPA$Sex=="Female"]<-0
baselineMVPA$Gender[baselineMVPA$Sex=="Male"]<-1
baselineMVPA$Gender<-as.numeric(baselineMVPA$Gender)
baselineMVPA$Gender<-scale(baselineMVPA$Gender, scale=FALSE)
baselineMVPA$School[baselineMVPA$Education=="High School Graduate" | baselineMVPA$Education=="High Scool Graduate"]<-1
baselineMVPA$School[baselineMVPA$Education=="Some College" | baselineMVPA$Education=="Associates Degree"]<-2
baselineMVPA$School[baselineMVPA$Education=="Postgraduate degree" | baselineMVPA$Education=="Bachelors degree"]<-3
baselineMVPA$School<-as.numeric(baselineMVPA$School)
baselineMVPA$School<-scale(baselineMVPA$School, scale=FALSE)

xsectional<-baselineMVPA[c(1,2,10:16)]

write.xlsx(xsectional, "xsectional.xlsx")


####Longitudinal Data###

longitudinalMVPA<-wide.data2[c(1,3,4,5,6,7,9,11,15,26)]

#Create slope and intercepts for time-varying variables#
MVPAslope<-lmer(MVPAMin~timefactor + (1+ time|ID), data3)
summary(MVPAslope)
MVPAslopes<-coef(MVPAslope)$ID
library(plyr)
MVPAslopes<-rename(MVPAslopes,c("time"="MVPASlope", "(Intercept)"="MVPAIntercept"))
MVPAslopes<-cbind(ID,MVPAslopes)
MVPAslopes<-MVPAslopes[c(1,2)]
detach("package:plyr", unload = TRUE)

SBslope<-lmer(SBMin~timefactor + (1+ time|ID), data3)
summary(SBslope)
SBslopes<-coef(SBslope)$ID
library(plyr)
SBslopes<-rename(SBslopes,c("time"="SBSlope", "(Intercept)"="SBIntercept"))
SBslopes<-cbind(ID,SBslopes)
SBslopes<-SBslopes[c(1,2)]
detach("package:plyr", unload = TRUE)

Slopes<-merge(MVPAslopes,SBslopes,by="ID")
longitudinal<-merge(longitudinalMVPA,Slopes,by="ID")

longitudinal$Groupscale<-scale(longitudinal$Treatment.Group, scale=FALSE)
longitudinal$Agescale<-scale(longitudinal$Age, scale=FALSE)
longitudinal$BMIscale<-scale(longitudinal$BMIbaseline, scale=FALSE)
longitudinal$MOCAscale<-scale(longitudinal$MoCA, scale=FALSE)
longitudinal$Gender[longitudinal$Sex=="Female"]<-0
longitudinal$Gender[longitudinal$Sex=="Male"]<-1
longitudinal$School[longitudinal$Education=="High School Graduate" | longitudinal$Education=="High Scool Graduate"]<-1
longitudinal$School[longitudinal$Education=="Some College" | longitudinal$Education=="Associates Degree"]<-2
longitudinal$School[longitudinal$Education=="Postgraduate degree" | longitudinal$Education=="Bachelors degree"]<-3
longitudinal$MVPASlopescale<-scale(longitudinal$MVPASlope, scale=FALSE)
longitudinal$SBSlopescale<-scale(longitudinal$SBSlope, scale=FALSE)

longitudinal.final<-longitudinal[c(1,2,13:18)]

write.xlsx(longitudinal.final, "longitudinalMVPA.xlsx")
```

#5:Additional reviewer requested analyses

##5.1: Does pain subscale changes relate to changes in MVPA

Data management for full dataset
```{r}
fullwide.data$time<-1
fullwide.data2<-fullwide.data[c(1,85,3:8,12,16,27,31,35,46,50,54,65,69,73,84)]
fulldata2 <- reshape(as.data.frame(fullwide.data2),idvar="ID",varying=c(9:20),direction="long",sep=".") 

fullID<-c("MO-001", "MO-006", "MO-009", "MO-011", "MO-012", "MO-023", "MO-028", "MO-030", "MO-033", "MO-035", "MO-056", "MO-063",
          "MO-072", "MO-077", "MO-078", "MO-079", "MO-080", "MO-081", "MO-082", "MO-086", "MO-092", "MO-095", "MO-101", "MO-104",
          "MO-106", "MO-108", "MO-109", "MO-113", "MO-120", "MO-126", "MO-127", "MO-128", "MO-137", "MO-138", "MO-141", "MO-142",
          "MO-151", "MO-153", "MO-154", "MO-157", "MO-159", "MO-167", "MO-170", "MO-185", "MO-188", "MO-190", "MO-193", "MO-196",
          "MO-203", "MO-205", "MO-209", "MO-218", "MO-219", "MO-221", "MO-223", "MO-238", "MO-253", "MO-255", "MO-256", "MO-259",
          "MO-276")
```

Create slope and intercepts for time-varying variables for full data
```{r}
fullMVPAslope<-lmer(MVPA_Min~factor(time) + (1+ time|ID) + Treatment.Group, fulldata2)
summary(fullMVPAslope)
fullMVPAslopes_coef<-coef(fullMVPAslope)$ID
library(plyr)
fullMVPAslopes_coef<-rename(fullMVPAslopes_coef,c("time"="MVPASlope", "(Intercept)"="MVPAIntercept"))
MVPAslopesframe<-cbind(fullID,fullMVPAslopes_coef)
MVPAslopesframe<-MVPAslopesframe[c(1,2)]
detach("package:plyr", unload = TRUE)

fullSBslope<-lmer(SB_Min~factor(time) + (1+ time|ID) + Treatment.Group, fulldata2)
summary(fullSBslope)
fullSBslopes_coef<-coef(fullSBslope)$ID
library(plyr)
fullSBslopes_coef<-rename(fullSBslopes_coef,c("time"="SBSlope", "(Intercept)"="SBIntercept"))
SBslopesframe<-cbind(fullID,fullSBslopes_coef)
SBslopesframe<-SBslopesframe[c(1,2)]
detach("package:plyr", unload = TRUE)

fullKOOSslope<-lmer(KOOS_Pain~factor(time) + (1+ time|ID) + Treatment.Group, fulldata2)
summary(fullKOOSslope)
fullKOOSslope_coef<-coef(fullKOOSslope)$ID
library(plyr)
fullKOOSslope_coef<-rename(fullKOOSslope_coef,c("time"="KOOSSlope", "(Intercept)"="KOOSIntercept"))
KOOSslopeframe<-cbind(fullID,fullKOOSslope_coef)
KOOSslopeframe<-KOOSslopeframe[c(1,2)]
detach("package:plyr", unload = TRUE)

Slopes1<-merge(MVPAslopesframe,SBslopesframe, by="fullID")
Slopes2<-merge(Slopes1,KOOSslopeframe, by="fullID")
```

Correlation for full data
```{r}
cor.test(Slopes2$KOOSSlope, Slopes2$MVPASlope)
cor.test(Slopes2$KOOSSlope, Slopes2$SBSlope)
```

Create slope and intercepts for time-varying variables for MRI data only
```{r}
slopeMVPA<-lmer(MVPA_Min~factor(Timepoint) + (1+ Timepoint|ID) + Treatment.Group, data)
summary(slopeMVPA)
slopeMVPA_coef<-coef(slopeMVPA)$ID
library(plyr)
slopeMVPA_coef<-rename(slopeMVPA_coef,c("Timepoint"="MVPASlope", "(Intercept)"="MVPAIntercept"))
slopeMVPAframe<-cbind(ID,slopeMVPA_coef)
slopeMVPAframe<-slopeMVPAframe[c(1,2)]
detach("package:plyr", unload = TRUE)

slopeSB<-lmer(SB_Min~factor(Timepoint) + (1+ Timepoint|ID) + Treatment.Group, data)
summary(slopeSB)
slopeSB_coef<-coef(slopeSB)$ID
library(plyr)
slopeSB_coef<-rename(slopeSB_coef,c("Timepoint"="SBSlope", "(Intercept)"="SBIntercept"))
slopeSBframe<-cbind(ID,slopeSB_coef)
slopeSBframe<-slopeSBframe[c(1,2)]
detach("package:plyr", unload = TRUE)

slopeKOOS<-lmer(KOOS_Pain~factor(Timepoint) + (1+ Timepoint|ID) + Treatment.Group, data)
summary(slopeKOOS)
slopeKOOS_coef<-coef(slopeKOOS)$ID
library(plyr)
slopeKOOS_coef<-rename(slopeKOOS_coef,c("Timepoint"="KOOSSlope", "(Intercept)"="KOOSIntercept"))
slopeKOOSframe<-cbind(ID,slopeKOOS_coef)
slopeKOOSframe<-slopeKOOSframe[c(1,2)]
detach("package:plyr", unload = TRUE)

Slopes3<-merge(slopeMVPAframe,slopeSBframe, by="ID")
Slopes4<-merge(Slopes3,slopeKOOSframe, by="ID")
```

Correlation for full data
```{r}
cor.test(Slopes4$KOOSSlope, Slopes4$MVPASlope)
cor.test(Slopes4$KOOSSlope, Slopes4$SBSlope)
```


#5.2: Were there any lasting effects of the intervention on MVPA?

Pre-specified contrasts
```{r}
#First contrast: Immediate versus delayed from baseline to week 8
c1=matrix(c(0,1,0,-1,0,0,0,0))
c2=matrix(c(1,0,-1,0,0,0,0,0))
c1st=c1-c2

#Second contrast: Delayed after versus delayed before intervention (long-term)
c1=matrix(c(1,0,-1,0,0,0,0,0))
c2=matrix(c(0,0,1,0,0,0,-1,0))
c2nd=c2-c1

#Third contrast: 
c1=matrix(c(0,1,0,-1,0,0,0,0))
c2=matrix(c(0,0,1,0,0,0,-1,0))
c3rd=c1-c2

#Fourth contrast
c1=matrix(c(0,1,0,-1,0,0,0,0))
c2=matrix(c(1,0,0,0,-1,0,0,0))
c4th=c1-c2

#Fifth contrast
c1=matrix(c(0,0,0,1,0,-1,0,0))
c2=matrix(c(0,0,0,0,1,0,-1,0))
c5th=c1+c2/2
```

Contrasts for full data
```{r}
fulldata2$timefactor<-as.factor(fulldata2$time)
MVPAfulllmer<-lmer(MVPA_Min~timefactor*Treatment.Group + (time|ID), fulldata2)
mdlMVPA.lsm<-lsmeans(MVPAfulllmer, ~Treatment.Group|timefactor)
contrast(mdlMVPA.lsm,list(c1st,c2nd,c3rd,c4th,c5th),by=NULL)
```

Contrasts for MRI sub-analysis
```{r}
data$timefactor<-as.factor(data$Timepoint)
MVPAsublmer<-lmer(MVPA_Min~timefactor*Treatment.Group + (Timepoint|ID), data)
mdlMVPAsub.lsm<-lsmeans(MVPAsublmer, ~Treatment.Group|timefactor)
contrast(mdlMVPAsub.lsm,list(c1st,c2nd,c3rd,c4th,c5th),by=NULL)
```


#5.3: Determine intervention-associated changes in MVPA and include in fMRI analysis
```{r}
library(dplyr)
MVPAchange<-wide.data2 %>%
  mutate(
    MVPAchange = case_when(
      Treatment.Group == "1" ~ MVPAMin.2 - MVPAMinbaseline,
      Treatment.Group == "2" ~ MVPAMin.3 - MVPAMin.2
    ),
    SBchange = case_when(
      Treatment.Group == "1" ~ SBMinbaseline - SBMin.2,
      Treatment.Group == "2" ~ SBMin.2 - SBMin.3
    ),
    MVPAchangescale = scale(MVPAchange, center = TRUE, scale = FALSE),
    SBchangescale = scale(SBchange, center = TRUE, scale = FALSE)
  )

MVPAchange<-MVPAchange[c(1,83,84)]
print(MVPAchange)

```


